




Import packages
```{r}
# install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyr")
install.packages("scales")


library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
```

Load data
```{r}
df <- read.csv("outputs/tax_share_bydecile.csv")

df_long <- df %>%
  pivot_longer(
    cols = c(vatable_share_exp, exempt_share_exp),
    names_to = "type",
    values_to = "share"
  ) %>%
  mutate(
    # type = factor(type, levels = c("vatable_share_exp", "exempt_share_exp")),
    share_prop = share / 100,   # Convert 56.12 → 0.5612
    NPCINC = factor(NPCINC, levels = sort(unique(NPCINC), decreasing = TRUE)),
    label = paste0(round(share, 1), "%")   # keep label in percent
  )

df_long <- df_long %>%
  mutate(
    type = factor(type, levels = c("vatable_share_exp", "exempt_share_exp"))
  )

vatable_sh <- ggplot(df_long, aes(x = NPCINC, y = share, fill = type)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(
    aes(label = label),
    position = position_fill(vjust = 0.5),
    color = "black",
    size = 6
  ) +
    coord_flip() +
    scale_y_continuous(labels = scales::percent_format()) +
    scale_fill_manual(
    values = c("vatable_share_exp" = "lightblue", "exempt_share_exp" = "orange"),
    labels = c("vatable_share_exp" = "Vatable", "exempt_share_exp" = "Exempt"),
    name = "",
    guide = guide_legend(reverse = TRUE)   # <-- FIX
  ) +
    labs(
    x = "Per Capita Income Decile",
    y = "Share of Total Expenditure"
  ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top")

ggsave(
  filename = "outputs/vatable_sh.png",
  plot = vatable_sh,
  width = 10,
  height = 6,
  units = "in",
  dpi = 500
)


